Title here

### 2020-09-23 analysis/paper/paper.tex
fraft
llegamos a la conclusion!

@@

We defined the Southern Annular Mode (SAM) as the leading EOF of the
monthly anomalies of geopotential field at 700 hPa south of 20\degree S
(citation?). The EOF was performed by computing the Singular Value
Decomposition of the data matrix consisting in 481 rows and 4176 columns
(144 points of longitude and 29 points of latitude). The values where
weighted by the square root of the cosine of latitude to account for the
non-equal area of each gridpoint \citep{chung1999}. This same method was
used at the rest of the levels considered in this paper.

To separate between the zonally symmetric and asymmetric components of
the SAM, we computed the zonal mean and anomalies of the full SAM
spatial pattern. The results are shown in Figure~\ref{fig:method} for
700hPa. The full spatial signal (\(\mathrm{EOF_1}(\lambda, \phi)\)) is
the sum of the zonally asymmetric (\(\mathrm{EOF_1^*}(\lambda, \phi)\))
and symmetric (\([\mathrm{EOF_1}](\lambda, \phi)\)) components. We then
compute the ``Full'', ``Asymmetric'' and ``Symmetric'' indices, by
regressing each geopotential field on these patterns (weighting by the
cosine of latitude).

The three indices are normalised by dividing them by the standard
deviation of the ``Full'' index at each level. This means that comparing
the magnitude between indices is meaningful, but it also means that not
every index will have unit standard deviation.

@@

We used monthly geopotential height at 2.5 longitude by 2.5 latitude
resolution from ERA5 \citep{hersbach2020} for the period 1979 to 2018.

Monthly temperature NOAA Global Surface Temperature (NOAAGlobalTemp) 5.0
degree latitude x 5.0 degree longitude global grid
\citep{vose2012, smith2008}. The same analysis was carried out using
CRUTEM4 \citep{osborn2014} (not shown).

We used monthly precipitation data from CPC Merged Analysis of
Precipitation \citep{xie1997} 2.5 degree latitude x 2.5 degree
longitude. CPCC: {[}schneider2015{]} \#FIXME

@@

Explain multiple regression interpretation \#FIXME.

We adjusted p-values for False Detection Rate following
\citet{wilks2016}.

@@

Figure \ref{fig:asymsam-timeseries} shows the Asymmetric and Symmetric
time series corresponding to 700 and 50hPa. At first glance the series
can be distinguished by their distributions. Whereas the tropospheric
incides are aproximately normally distributed, the stratospheric indices
are more long-tailed; that is, extreme values (both negative and
positive) abound. The Asymmetric series have both more variability in
the higher frequencies than the Symmetric series.

The stratospheric Symmetric SAM varies strongly witha two-year period,
which can be seen using spectral methods (Figure A3) or in the
autocorrelation structure (Figure A4). There is a local peak at 2 years
in the periodigram of the tropospheric Symmetric SAM also, although it's
not statistically significant. In the troposphere the most significant
peak of variability is found in the Asymmetric index at around 3.6
months.

From Figure \ref{fig:asymsam-timeseries} we can see that the Asymemtric
and Symmetric time series appear to be correlated. Moreover, looking at
the extremes in the stratosphere, the Symmetric serie appears to lag the
Asymmetric series (see, for example, the positive events on late 1987
marked with a circle). We show these correlations, across all the levels
of the reanalysis and for zero and -1 lag (Asymmetric index leading the
Symmetric index), in Figure \ref{fig:cor-lev}.

@@

Zero-lag correlations between the Asymmetric and Symmetric series are
relatively constant throught the troposphere, fluctiating between 0.39
and 0.45. One-month-lag correlations are similarly constant but
significantly reduced, hovering around 0.17. In the stratosphere,
zero-lag correlations drop to a minimum of 0.21 at 20 hPa and then it
increases again monotonically with height up to the uppermost level of
the reanalysis. At the same time, one-month-lag correlations increase
with height.

@@

Figure \ref{fig:cross-correlation}a) shows (zero-lag) cross-correlation
across levels for the Full, Symmetric and Asymmetric SAM indices. For
the Full SAM (panel a), high values below 100 hPa reflect the vertical
(zero-lag) coherency throughout the troposfere. Above 100 hPa
correlation between levels falls off more rapidly, indicating less
coherent (zero-lag) variability. Still there is a non negligible
correlation between the troposphere and the lower-to-middle
stratosphere. Examining panels b and c, we see that the Asymemtric and
Symmetric SAM share the same high level of coherency in the troposphere
but they differ in their stratospheric behaviour. As evidenced by the
wider dark red areas near the diagonal in Figure
\ref{fig:cross-correlation}b) vs.~Figure \ref{fig:cross-correlation}c),
stratospheric coherency is stronger for the Asymmetric SAM than the
Symemtric SAM. The stratospheric Symmetric SAM seems to connect more
strongly to the trosposphere than the Asymmetric SAM; this can be seen
by the lower correlation values in the top right left of Figure
\ref{fig:cross-correlation}b) in comparison with Figure
\ref{fig:cross-correlation}c).

@@

Figure \ref{fig:trends} shows normalised decadal trends for each index
for the whole period 1979-2918 along with the 95\% confidence interval
in shading for the whole year (row a) and separed by trimesters (rows b
through e). As previously documented (e.g. \citet{fogt2020}), there is a
statistically significant increase towards more positive SAM (panel
a.1), which is evident only in Summer and Autumn (panels b.1 and c.1).
We observe these increases mainly in the troposphere, reaching their
maximum at at 100 hPa in Summer. By separating the SAM signal in its
Asymmetric and Symmetric parts, we can not only see that these trends
are almost entirely due to the Symmetric component (columns 2
vs.~columns 3), but in some cases the trends become more clear. In
Summer, the Asymmetric SAM has a statistically non significant negative
trend in the middle troposphere that obscures the signal; as a result,
trends computed using only the Symmetric component are more clear
(compare the shading region in panel b.1 and b.3). In Autumn, using the
Symmetric SAM reveals a statistically significant positive trend in the
stratosphere that is not significant using the Full index.

@@

\begin{figure*}
\includegraphics{2d-regr-1} \caption[Regression patterns of geopotential height at 30, 300 and 700 hPa with the Full, Asymmetric and Symmetric SAM]{Regression patterns of geopotential height at 30, 300 and 700 hPa with the Full, Asymmetric and Symmetric SAM. The regression patterns for Asymmetric and Symmetric SAM are the result of one multiple regression using both indices, not of two simple regressions involving each index by itsef. Points marked on panel b.2 are the location of the reference points used by \cite{raphael04} for its Zonal Wave 3 index.}\label{fig:2d-regr}
\end{figure*}

@@

In the stratosphere, the spatial pattern associated with the Full SAM is
more clearly dominated by a zonally symmetric, monopolar structure
(panel a.1) which is, however, not perfectly centered in the south pole.
The monopole obtained by multiple regression with the Asymmetric and
Symmetric SAM (panel c.1) is much more symmetric and the shift from
total symmetry is captured by the regression pattern of the Asymmetric
SAM as a wave-1 with maximum anomalies above the Belinghausen Sea on the
Western Hemisphere and and Davids Sea in the Eastern Hemisphere (panel
b.1).

In the troposphere, panel a.2 shows the well known zonally symmetrical
annular mode \emph{contaminated} with zonal asymmetries in the form of a
wave-3. The regression using the Asymmetric and Symmetric SAM indices
successfully disentangle both structures. The Asymmetric component gives
rise to a cleaner zonal wave (panel b.2) and the Symemtric component is
assosiated with an trully annular mode, almost devoid of zonal
asymmetries (panel c.2). The wave-3 pattern observed in panel b.2 is
rotated by half a wavelength from the average position of the mean
wave-3 pattern asociated with \citet{raphael2004}'s ZW3 index, whose
reference locations are marked with points in the figure. Thus, the
tropospheric Asymmetric SAM index responds to zonal displacements in the
position of the wave-3 pattern.

@@

The amplitude of each zonal wave number at each latitude at 50 hPa and
700 hPa is shown in Figure \ref{fig:wave-amplitude}, where wave number
zero represents the amplitude of the zonal mean. Comparing between rows,
this Figure quantifies the relatively clean separation between the
zonally symmetric and zonally asymmetric structures, as its evident how
the mixture of waves of the Full field (column a) is very similar to the
sum of the waves of the Asymmetric and Symmetric field (columns b and c,
respectively). Column b of Figure \ref{fig:wave-amplitude} shows that
the Asymmetric SAM is overwhelmingly dominated by wave 1 in the
stratosphere (panel b.1), while in the troposphere it is composed of
zonal waves 3 to 1 in decreasing level of importance (panel b.2).

@@

To analyse the vertical structure of the geopotential anomalies
asociated with the asymetric SAM index, we show a vertical cross section
of regressions of mean geopotential height between 65ªS and 40ªS for the
50 hPa Asymmetric SAM index (panel a) and for the 700 hPa Asymmetric SAM
index (panel b) (Figure \ref{fig:vertical-regression}).

The geopotential anomalies associated with the stratospheric SAM (panel
a) are clearly constrained to the stratosphere, which underscores the
disconnect between the stratospheric and tropospheric Asymmetric SAM.
The vertical structure this signal tilts about 60ª to the West between
100 hPa and 1 hPa, suggesting baroclinic processes and polarward
transport of heat (\#FIXME is this ok?). Interestingly, the signal in
the stratosphere maximises near 10 hPa despite using the 50 hPa index
for the regression.

@@

The structures shown in panels a and b in Figure
\ref{fig:vertical-regression} are surprisignly robust to the choice of
index level. For any stratospheric (above 100 hPa) index, the resulting
anomalies are very similar to the wave-1 structure with maximum near 10
hPa in panel a. Conversely, for any tropospheric (below 100 hPa) index,
the result is very similar to panel b. Pattern cross-correlation is
greater than 0.9 within stratospheric index levels and tropospheric
index levels and less than .4 accross stratospheric and tropospheric
levels (Figure A8), i.e.~regressions obtained using Asymmetric SAM
indices in the troposphere are very similar to each other and different
to the ones obtaines using indices in the stratosphere and vice versa.
The patterns mainly change in amplitude. The tropospheric pattern is
maximised by the 300 hPa Asymmetric SAM index and the stratospheric
pattern increased monotonically with height.

The wave-3 pattern from Figure \ref{fig:2d-regr} panel b.2 is very
similar to the teleconnection pattern associated with the ENSO. Indeed,
\citet{fogt2011} showed that there is a significant relationship between
the SAM and the ENSO. The correlation between the full SAM and the ENSO
as measured by the Multivariate Enso Index \citep{wolter2011} is -0.19.
This relationship is captured entirely the Asymmetric SAM, as this index
has a partial correlation of -0.27 with the MEI, whereas the Symmetric
SAM's partial correlation with the MEI is essentially null
(\ensuremath{-7.9\times 10^{-4}}).

@@

\subsubsection{Temperature}

@@

Figure \ref{fig:regr-air-season} shows regression coefficients of each
index at 700 hPa with surface temperature for each trimester. It is
evident that the Asymmetric and Symmetric SAM indices are associated
with overall distinct temperature patterns which can be obscured when
using the Full SAM index. The Symmetric SAM signal is weaker than the
Asymmetric SAM, as evidenced by the relatively smaller and les
sstatistically significant regression coefficients in row 3 of Figure
\ref{fig:regr-air-season} compared with row 2.

In DJF (row 1), the strong negative signal in the tropical Pacific in
panel a.1 is mostly associated with the Asymmetric component (panel
b.1), as is it largely absent in the Symmetric component (panel c.1).
Furthermore, the Asymmetric SAM is also asociated with low temperature
anomalies in the Indian ocean, but this signal is obscured by the
Symmetric variability and thus lost in the Full SAM. Over the
continents, the Asymmetric SAM is assoiated with negative temperature
anomalies which, again, mostly disappear in the Full SAM regression.
Temperature regression with the Symmetric SAM (panel c.1) shows a ring
of negative anomalies around Antarctica surrounded by positive
anomalies. This pattern of anomalies is consistent with thermal wind
balance and the intesification and pole-ward migration of the westerlies
commonly linked to the SAM.

The patterns seen in MAM and JJA (rows 2 and 3) are not robustly
significant in the sense that there are no areas with p-values below
0.05 when controlling for FDR following \citet{wilks2016}. Nevertheless,
it is interesting to note that in both trimesters, the sign of the
regression is consistently flipped between the Asymmetric and Symmetric
regressions. In South America, for example, the Asymmetric SAM is
associated with positive temperature anomalies in MAM and negative
temperature anomales in JJA, while the oposite is the case for the
Symemtric SAM.

Finally, in SON (row 4) there is no significant temperature signal
associated with the Symmetric SAM (panel c.4), while the Asymmetric SAM
shows a relatively robust signal in the equatorial Pacitic, Australia,
and even Southeast South America.

\subsubsection{Precipitation}

Regression of the SAM indicies with seasonal mean precipitation are
shown Figures \ref{fig:pp-regr-oceania} and \ref{fig:pp-regr-america}
for Australia and New Zealand, and South America respectively. (We
didn't detect any significant signal in South Africa.)

@@

In Australia (Figure \ref{fig:pp-regr-oceania}), the annual-level
regression shows that the Full SAM is associated with a statistically
significant increase in precipitation in the Southeastern region (panel
a.1), which reproduces the results from \citet{gillett2006}. The
separation between Asymmetric and Symmetric SAM suggest that this
increase is explained by the Symmetric SAM only in the East coast (panel
c.1), which is consistent with the increased easterly flow clearly seen
in relation with this index. The Asymmetric SAM appears related to
increased precipitation in the West coast of Southeastern Australia
(panel b.2), explained by the anomalous \emph{westerly} circulation
transporting moist air to the continent.

The seasonal-level regressions show statistically significant anomalies
only in SON, with a pattern similar to the annual-level regression
(panel a.5). Panels b.5 and c.5 don't show a clear separation between
the Asymmetric and Symmetric SAM. If anything, the positive and more
significant regression coefficients in panel b.5 vs pane c.5 would
suggest more influence of the Asymmetric than the Symmetric SAM, going
against the interpretation gathered from the annual-level regressions.
This Spring signal is broadly consistent with \citet{hendon2007}, but
whereas \citet{hendon2007} also detected a strong signal in Summer,
panel a.2 shows no statistically significant association (although the
coeffcients have the consistent sign).

@@

In South America (Figure \ref{fig:pp-regr-america}), the annual-level
regression shows that the SAM is associated with statistically
significant precipitation decrease in Southeastern South America (SESA)
and Southern Chile and non-significant increase in South Brazil, near
the South Atlantic Convergence Zone (SACZ) (panel a.1).

Panels b.1 and c.1 show a remarkably clean separation between the
Asymmeric SAM --associated with the Southeastern South American and
Southern Brazilian signals-- and the Symmetric SAM --associated with the
signal in Southern Chile. This separation is consistent with the
mechanisms responsible for these effects. In Southern Chile, the reduced
westerly flow reduce moisture transport from the Pacific Ocean (cita??
\#FIXME).

In Southeastern South America, anomalous meridional winds lead to less
precipitation by inhibiting moisture convergence from the South American
Low Level Jet \citep{silvestri2009}. The increased precipiation in the
South Atlantic Convergence Zone, on the other hand, appears to be
related to modulation of the anomalous SAM circulation in SACZ events
which lead to more frequent and intense SACZ events during positive SAM
\citep{rosso2018}.

@@

Except during Winter, the seasonal-level regression all show these
similar patterns although not as cleanly and only in some cases
statistically significant.

\subsubsection{Sea ice}

\begin{figure*}
\includegraphics{regr-ice-1} \caption[Seasonal regression of SAM indices with sea ice concentration]{Seasonal regression of SAM indices with sea ice concentration. \#FIXME}\label{fig:regr-ice}
\end{figure*}

>>>>

To describe the Southern Annular Mode and its variability we used
monthly geopotential height at 2.5\degree longitude by
2.5\degree latitude of horizontal resolution and 37 vertical isobaric
levels from ERA5 \citep{hersbach2020} for the period 1979 to 2018. We
restrict our analysis to the post-satellite era to avoid any confounding
factors arising from the introduction of satellite observations.

We describe the relationship between the SAM indices and temperature and
precipitation. We use temperature data from NOAA's Merged Land Ocean
Global Surface Temperature Analysis V4.0.1 \citep{vose2012, smith2008},
which blends land and ocean temperature analysis into a monthly global
grid 5\degree lagitude by 5\degree longitude. For precipitation, we use
monthly, 0.5\degree latitude by 0.5\degree longitude data from the
Global Precipitation Climatology Centre \citep{schneider2015}.

@@

Traditionally the Souther Annular Mode (SAM) is defined as de leading
Empiral Orthogonal Mode (EOF) of sea level pressure or geopotential
height at lower levels \citep{ho2012}. Following \citet{baldwin2001}, we
extend that definition vertically and use the term SAM to refer to the
the leading EOF of the monthly anomalies of geopotential field south of
20\degree S at each level. We performed EOFs by computing the Singular
Value Decomposition of the data matrix consisting in 481 rows and 4176
columns (144 points of longitude and 29 points of latitude). We weighted
the values by the square root of the cosine of latitude to account for
the non-equal area of each gridpoint \citep{chung1999}.

To separate between the zonally symmetric and asymmetric components of
the SAM, we computed the zonal mean and anomalies of the full SAM
spatial pattern, as shown in Figure~\ref{fig:method} for 700hPa. The
full spatial signal (\(\mathrm{EOF_1}(\lambda, \phi)\)) is the sum of
the zonally asymmetric (\(\mathrm{EOF_1^*}(\lambda, \phi)\)) and
symmetric (\([\mathrm{EOF_1}](\lambda, \phi)\)) components. We then
compute the ``Full'', ``Asymmetric'' and ``Symmetric'' indices, by
regressing each geopotential field on the respective patterns (weighting
by the cosine of latitude). The three indices are normalised by dividing
them by the standard deviation of the ``Full'' index at each level. As a
result, the magnitude between indices is comparable. However, only
``Full'' index will have unit standard deviation per definition.

@@

Our method assumes linearity in the asymmetric component of the SAM.
That is, we assume that zonal symmetries associated with positive SAM
are oposite and equal to the ones associated with negatie SAM.
\citet{fogt2012}'s composites (their Figure 4) suggest that this might
not be entirely valid, although we argue that much of that apparent
non-linearity is due to the heterogenous nature of the selected years
for constructing the composites. Using our data (from 1979 to 2018),
seasonal composites of zonal anomalies of 700 hPa geopotential height
for SAM+ (Full SAM index greater than 1 standard deviation) and SAM-
(smaller than negative 1 standard deviation) show relatively high
pattern correlations all seasons and are visually very linear (Figure
A9). Therefore, we belive that our method is at the very least a
reasonable approximation of the phenomenon.

By computing a single EOF pattern using data for all months we are
assuming that the zonal anomalies of the SAM are the same in all
seasons. Geopotential zonal anomalies computed by projecting the first
EOF \emph{of each season} are very similar to each other (Figure A10)
and show pattern correlations between 0.65 (DJF with JJA) and 0.9
(between MAM and SON). Based on this, we believe that our initial
assumption is not unreasonable.

Finally, we assume that the zonally asymmetric pattern is stationary in
time. \citet{silvestri2009} suggest that this might not be the case
between 1958 and 2004 but the period we analyse is much shorter
(1979-2018) so it's unlikely that we could observe significant changes.
Moreover, zonal asymmetry of the spatial patterns for the two halves of
the period (1979 to 1998 and 1999 to 2018) show no systematic change
(Figure A11).

@@

We perform linear regression to quantify the association between the SAM
indices and other variables. Since the Asymmetric and Symmetric SAM
indices are significantly correlated with each other, to capture the
variability explained uniquely by each index we use one multiple linear
regression instead of two simple linear regressions. To obtain the
linear coefficients of a variable \(X\) (geopotential, temperature,
precipitation, etc\ldots{}) with the Asymmetric SAM (\(SAM_a\)) and
Symmetric SAM (\(SAM_s\)) we fit the ecuation

\[
X(\lambda, \phi, t) = \alpha(\lambda, \phi) SAM_a + \beta(\lambda, \phi) SAM_s + X_0(\lambda, \phi) +  \epsilon(\lambda, \phi, t)
\]

where \(\lambda\) and \(\phi\) are the longitude and latitude, \(t\) is
the time, \(\alpha\) and \(\beta\) are the linear coefficients, \(X_0\)
and \(\epsilon\) are the constant and error terms. From this equarion,
\(\alpha\) represents the (linear) association of \(X\) with the
variability of the Asymmetric SAM that is not explained by the
variability of the Symmetric SAM; in other words, it is proportional to
the partial correlation of \(X\) and the Asymmetric SAM, controlling for
the effect of the Symmetric SAM and viceversa for \(\beta\).

At 2.5\degree by 2.5\degree resolution, a single regression field is
composed of thousands of regressions. In such case, using naive p-values
to test for significance leads to misleading results
\citep{walker1914, katz1991}. While there are multiple proposed
solutions in the literature, \citet{wilks2016} suggests that adjunting
p-values by controlling for the False Discovery Rate
\citep{benjamini1995} is a simple and effective method to ameliorate
this issue. Therefore, p-values showed in regression fields are all
adjusted following \citet{benjamini1995}.

When performing a separate regression for each trimester (DJF, MAM, JJA,
SON) we first average the relevant variables to obtain a single value
for each year and each trimester.

@@

The temporal evolution of the Assymmetric and Symmetric SAM was firstly
asssesed. Figure \ref{fig:asymsam-timeseries} shows the corresponding
time series for 700 hPa and 50 hPa and their corresponding density
estimates. We selected these two levels as representative of the
tropospheric and stratospheric variability respectively. As will be
shown later, both indices are highly coherent within each atmospheric
layer, therefore is reasonable to take one level as representative of
each layer.

Month-to-month variability is evident for both indices, with noisy
variations in the low frequency. At first glance the series can be
distinguished by their distributions. Compared to the stratospheric
indices, the stratospheric indices are much more long-tailed; that is,
extreme values (both negative and positive) abound. The Asymmetric
series have both more variability in the higher frequencies than the
Symmetric series.

The stratospheric Symmetric SAM varies strongly with a two-year period,
which can be seen by spectral analysis (Figure A3). This might suggests
a link between stratospheric SAM variability and QBO. There is a local
peak at 2 years in the periodigram of the tropospheric Symmetric SAM
also, although it's not statistically significant. In the troposphere
the most significant peak of variability is found in the Asymmetric
index at around 3.6 months.

From Figure \ref{fig:asymsam-timeseries} we can see that the Asymemtric
and Symmetric time series appear to be correlated. Moreover, looking at
the extremes in the stratosphere, the Symmetric serie appears to lag the
Asymmetric series (see, for example, the positive events on late 1987).
We show these correlations, across all the levels of the reanalysis for
zero and -1 lag (Asymmetric index leading the Symmetric index), in
Figure \ref{fig:cor-lev}. Zero-lag correlations between the Asymmetric
and Symmetric series are relatively constant throught the troposphere,
fluctiating between 0.39 and 0.45. One-month-lag correlations are
similarly constant but significantly reduced to around 0.17. In the
stratosphere, zero-lag correlations drop to a minimum of 0.21 at 20 hPa
and then it increases again monotonically with height up to the
uppermost level of the reanalysis (although results near the top of the
models are to be interpreted with care). At the same time, one-month-lag
correlations increase with height. As a consequence, statospheric
Asymmetric index tend to precede corresponding Symmetric index.

@@

@@

Figure \ref{fig:cross-correlation}a shows (zero-lag) cross-correlation
across levels for the Full, Symmetric and Asymmetric SAM indices. For
the Full SAM (panel a), high values below 100 hPa reflect the vertical
(zero-lag) coherency throughout the troposfere. Above 100 hPa
correlation between levels falls off more rapidly, indicating less
coherent (zero-lag) variability. Therefore, there is a non negligible
correlation between the troposphere and the lower-to-middle
stratosphere. Examining panels b and c, we see that the Asymemtric and
Symmetric SAM share the same high level of coherency in the troposphere
but they differ in their stratospheric behaviour. Stratospheric
coherency is stronger for the Asymmetric SAM than the Symemtric SAM. The
stratospheric Symmetric SAM seems to connect more strongly to the
trosposphere than the Asymmetric SAM.

@@

The trends for each of the indices (Full, Symmetric, Assymetric) were
evaluate for the whole period 1979-2018 at each level (Figure
\ref{fig:trends}) for the whole year and separed by trimesters. The Full
SAM index presents a statistically significant trend (panel a.1) that
extends throught the troposphere up to about 50 hPa and reaches its
maximum value at 100 hPa. The seasonal trends (rest of column a)
indicate that positive trends are present in Autumn and particularly in
in Summer, where the 100 hPa maximum is much more defined. Positive
trends have been documented by previous studies \citep[e.g.][ and
references therein]{fogt2020} using indices of the SAM based on surface
or near-surface circulation.

By separating the SAM signal in its Asymmetric and Symmetric parts, we
can not only see that these trends are almost entirely due to the
Symmetric component (colum b vs.~column c), but in some cases the trends
become more clear. In Summer, the Asymmetric SAM has a statistically non
significant negative trend in the middle troposphere that obscures the
trend in the Full index; as a result, trends computed using only the
Symmetric component are more clear (compare the shading region in panel
a.2 and c.2). In Autumn, using the Symmetric SAM reveals a statistically
significant positive trend in the stratosphere that is not significant
using the Full index.

@@

\begin{figure*}
\includegraphics{2d-regr-1} \caption[Regression patterns of geopotential height at 30, 300 and 700 hPa with the Full, Asymmetric and Symmetric SAM]{Regression patterns of geopotential height at 30, 300 and 700 hPa with the Full, Asymmetric and Symmetric SAM. The regression patterns for Asymmetric and Symmetric SAM are the result of one multiple regression using both indices, not of two simple regressions involving each index by itsef. Points marked on panel b.2 are the location of the reference points used by \cite{raphael2004} for their Zonal Wave 3 index.}\label{fig:2d-regr}
\end{figure*}

@@

In the stratosphere, the spatial pattern associated with the Full SAM is
more clearly dominated by a zonally symmetric, monopolar structure
(panel a.1) which is, however, not perfectly centered in the South Pole.
The monopole obtained by multiple regression with the Asymmetric and
Symmetric SAM (panel c.1) is much more symmetric and the shift from
total symmetry is captured by the regression pattern of the Asymmetric
SAM as a wave-1 with maximum anomalies above the Belinghausen Sea on the
Western Hemisphere and and Davids Sea in the Eastern Hemisphere (panel
b.1).

In the troposphere, panel a.2 shows the well known combination of
zonally symmetrical annular mode with zonal asymmetries in the form of a
wave-3. The regression using the Asymmetric and Symmetric SAM indices
successfully disentangle both structures. The Asymmetric component gives
rise to a cleaner zonal wave (panel b.2) and the Symemtric component is
assosiated with an trully annular mode, almost devoid of zonal
asymmetries (panel c.2). The wave-3 pattern observed in panel b.2 is
rotated by half a wavelength from the average position of the mean
wave-3 pattern asociated with \citet{raphael2004}'s ZW3 index, whose
reference locations are marked with points in the figure. Thus, the
tropospheric Asymmetric SAM index represents a zonal displacement in the
position of the climatological wave-3 pattern.

@@

The amplitude of first zonal wave numbers at each latitude at 50 hPa and
700 hPa is shown in Figure \ref{fig:wave-amplitude}, where wave number
zero represents the amplitude of the zonal mean. Column b shows that the
Asymmetric SAM is overwhelmingly dominated by wave 1 in the stratosphere
(panel b), while in the troposphere it is composed of zonal waves 3 to 1
in decreasing level of importance (panel b). Looking at panel b.2 from
Figure \ref{fig:2d-regr}, it becomes apparent that zonal wavess 1 and 2
modulate the amplitude of zonal wave 3, which --as mentioned before-- is
larger in the Western Hemisphere than in the Easten Hemisphere.

@@

To analyse the vertical structure of the geopotential anomalies
asociated with the asymetric SAM index, we show a vertical cross section
of regressions of mean geopotential height between 65\degree S and
40\degree S for the 50 hPa Asymmetric SAM index (panel a) and for the
700 hPa Asymmetric SAM index (panel b) (Figure
\ref{fig:vertical-regression}). The geopotential anomalies associated
with the stratospheric Asymmetric SAM (panel a) are clearly constrained
to the stratosphere, which underscores the uncoupling between the
stratospheric and tropospheric Asymmetric SAM. The vertical structure of
this signal tilts about 60\degree to the West between 100 hPa and 1 hPa,
suggesting baroclinic processes. Interestingly, the signal in the
stratosphere maximises near 10 hPa despite using the 50 hPa index for
the regression.

@@

The structures shown in panels a and b in Figure
\ref{fig:vertical-regression} are surprisignly robust to the choice of
index level. For any stratospheric (above 100 hPa) index, the resulting
anomalies are very similar to the wave-1 structure with maximum near 10
hPa in panel a. Conversely, for any tropospheric (below 100 hPa) index,
the result is very similar to panel b. The patterns mainly change in
amplitude.

The wave-3 pattern from Figure \ref{fig:2d-regr} panel b.2 is very
similar to the Pacific-South American Pattern \citep{mo1987, kidson1988}
which is a teleconnection pattern associated with the ENSO
\citep{karoly1989}. Indeed, \citet{fogt2011} showed that there is a
significant relationship between the SAM and the ENSO. The correlation
between the full SAM and the ENSO as measured by the Oceanic Niño Index
\citep{bamston1997} (ONI) is -0.16. This relationship is captured
entirely the Asymmetric SAM, as this index has a partial correlation of
-0.26 with the ONI controlling for the effect of the Symmetric SAM,
whereas the Symmetric SAM's partial correlation with the ONI is
essentially null (0.019). We performed the same analysis using the
Multivariate Enso Index \citep{wolter2011} and the Southern Oscillation
Index \citep{ropelewski1987} to conclude that these results do not
depend on the ENSO index used.

@@

@@

The SAM has been shown to be associated with important surface variables
such as temperature and precipitacion \citep[e.g.][and see
\citet{fogt2020} for a review]{gillett2006}. Naturally, most studies on
the surface impacts of the SAM are based on an index identical or
analogous to what we call Full SAM index (\citet{fogt2012} being the
only exception that we are aware of). We regress surface temperature and
precipitation onto each of the three SAM indices to see if there are
diferent surface impact associated with the asymmetric and symmetric SAM
circulation.

Figure \ref{fig:regr-air-season} shows regression coefficients of each
index at 700 hPa with surface temperature for each trimester. In Summer
positive values of the Full SAM index (panel a.1) are associated with
negative temperature anomalies near Antarctica which are surrounded by a
ring of positive anomalies. The ring is not zonally symmetric, as there
are three clear local maximums around 30\degree W, 15\degree E and
50\degree E and a local minimum (with negative sign) around
120\degree W. In the tropics, there are negative anomalies in the
equatorial Pacific, consistent with the negative correlation between SAM
and ENSO. Panels b.1 and c.1 show temperature anomalies associated with
positive values of the Asymmetric and Symmetric SAM, respectively. Both
the local maximums in the ring and the anomalies in the Pacific regions
are present mostly on the Asymetric SAM regression map, while
temperature patterns linked to positive Symmetric SAM show a more
zonally consistent ring and less relation to the tropics. Noticeable,
temperature anomalies in the Indian ocean, South Africa and Australia
are strongly related to positive values of Asymmetric SAM. This signal
is not present in the regression pattern with the Full SAM. Spring (row
4) features very similar patterns but of generally smaller in magnitude
and statistical significance.

In Autumn and Winter (rows 2 and 3) the postive ring is only present
through its local maximums in the regression with the Full SAM. There
are also negative anomalies in Southern Australia, and positive
anomalies over New Zealand and Southern South America. These patterns
are not significant in the sense that there are no areas with p-values
below 0.05 when controlling for FDR following \citet{wilks2016}.
However, repeating this analysis with 2-meter temperature from ERA5
resulted in similar patterns that were statistically significant.
Moreover, similar features were observed in station measurements by
\citet{jones2019}, although using data from 1957 to 2016.

The pattern of negative anomalies in the pole surrounded by positive
anomalies roughly seen in all seasons --although with varying intensity
and small-scale details-- is consistent with the intensification and
poleward migration of the westerlies commonly linked to the SAM. It's
then not surprising to see it more clearly in association with the
Symemtric SAM (at least in Summer and Spring).

These results suggests that Asymmetric and Symmetric SAM indices are
associated with overall distinct temperature patterns which may not be
apparent when using the Full SAM index.

Figure \ref{fig:regr-air-season} column b can be partially compared with
Figure 11 from \citet{fogt2012}. Although they used station data from
1958 to 2001, a lot of the characteristics are reproduced here, such as
the strong signal in New Zealand and Australia in Summer and Spring.

Regression of the SAM indicies with seasonal mean precipitation and 700
hPa geopotential height are shown Figures \ref{fig:pp-regr-oceania} and
\ref{fig:pp-regr-america} for Australasia and South America
respectively. South Africa is not shown because no significant signal
was detected there.

@@

In Australia, the annual regression shows that the Full SAM index is
positively associated with precipitation in the Southeastern region
(Figure \ref{fig:pp-regr-oceania} panel a.1), which reproduces the
results from \citet{gillett2006}. The separation between Asymmetric and
Symmetric SAM suggest that this positive anomaly is explained by the
Symmetric SAM only in the East coast (panel c.1). Geopotential anomalies
associated with this index (black contours) are indicative of easterly
flow from the Tasman Sea, which could explain the positive anomalies in
precipitation as found by \citet{hendon2007}. The Asymmetric SAM appears
related to increased precipitation in the West coast of Southeastern
Australia (panel b.2), which could similarly be explained by the
anomalous westerly circulation transporting moist air to the continent
from the Indian Ocean.

This Spring signal is broadly consistent with \citet{hendon2007}, but
whereas \citet{hendon2007} also detected a strong signal in Summer,
panel a.2 shows no statistically significant association (although the
coeffcients have the consistent sign).

The seasonal-level regressions show statistically significant anomalies
only in Spring, when positive Full SAM is associated with positive
precipitation anomalies in Eastern Australia (panel a.5). In this
trimester the Symmetric SAM seems to be associated with precipitation in
a relatively reduced area of the East Coast (panel c.5) while the
positive precipitation anomalies related with positive Asymmetric SAM
affect all Estern Australia (panel b.5).

In Summer, positive Full SAM index is associated with with positive
precipitation anomalies in Western and Eastern Australia, particularly
in the North East (panel a.2). The Eastern part being dominated by the
relationship with the Symmetric SAM and the Western, by the Asymemtric
SAM. In Autumn, the regression with Full SAM shows positive values in
the North, similar to Summer, and a broad area of positive values in the
North-East to South-West direction. This structure seems to be
associated with the Symmetric SAM, while the Northern positive values
are associated with the Asymmetric SAM. In Winter we see the same NE to
SW aligned anomaly (although with much reduced amplitude) that is also
present only in relation with the Symmetric SAM. None of these
regression coefficients are statistically signifincat at the 95\% level

@@

In South America (Figure \ref{fig:pp-regr-america}), the annual-level
regression shows that the SAM is associated with statistically
significant precipitation decrease in Southeastern South America (SESA)
and Southern Chile and non-significant increase in South Brazil, near
the South Atlantic Convergence Zone (SACZ) (panel a.1). Panels b.1 and
c.1 show a remarkably clean separation between the Asymmeric SAM
--associated with the Southeastern South American and Southern Brazilian
signals-- and the Symmetric SAM --associated with the signal in Southern
Chile.

Except Winter, seasonal-level regressions mirror this same pattern. Even
if not statistically significant, they all show negative values in
Southeastern South America and Southern Chile along with positive values
in Southern Brazil in relation with the Full SAM. The separation of
these features between the Asymmetric SAM and Symmetric SAM regression
maps is also rather consistent.

The anomalous circulation at 700 hPa associated with the Symmetric SAM
(panel c.1) indicate anomalous Easterly flow over Southern Chile. This
leads to reduced influx of moist air from the Pacific Ocean which, is
the main source of precipitable water in that region. On the other hand,
the anomalous circulation associated with positive values of Asymmetric
SAM (panel b.1) in the Atlantic is anticyclonic in the South and
cyclonic in the North. This creates anomalous South-Easterly flow over
Southeastern South America, which inhibits the flow of the Low Level Jet
to the region \citep[\citet{zamboni2010}]{silvestri2009}. This same
pattern was found to be associated with increased precipitation in
Southern Brazil during South Atlantic Convergence Zone events
\citep{rosso2018}.

@@
